\part{Foundations}

\chapter{Probabilistic Inference}

\begin{exercise}
\item Bayes rule for medical diagnosis
	\par\smallskip
	After your yearly checkup, the doctor has bad news and good news. The bad news is that you tested
	positive for a serious disease, and the test is \(99\%\) accurate (i.e., the probability of testing positive
	given that you have the disease is \(0.99\), as is the probability of testing negative given that you dont
	have the disease). The good news is that this rare disease, striking only one in \(10,000\) people.
	What are the chances that you actually have the disease?
	\par\smallskip
	The probability matrix will be given by \cref{ch2_ex1}.
	\begin{table}
	\centering
		\begin{tabular}{l | c c}
			& Y = 0 & Y = 1\\ \hline
			H = 0 & 0.99 & 0.01\\
			H = 1 & 0.01 & 0.99
		\end{tabular}
		\caption{The probability matrix for the given rare disease.}
		\label{ch2_ex1}
	\end{table}
	The probability of the prevalence of the disease is \(p(H = 1) = 1 / 10000\) for the population. The 
	probability that we are infected  and tested positive is
	\begin{align}
		p(H = 1\mid Y = 1) 
		&= \frac{p(Y = 1\mid H = 1) p(H = 1)}{p(Y = 1\mid H = 1) p(H = 1) + p(Y = 1\mid H = 0) p(H = 0)}
		\notag\\
		&= \frac{0.99\cdot 0.0001}{0.99\cdot 0.0001 + 0.01 \cdot (1 - 0.0001)}\notag\\
		&= 0.0098\text{ or } 0.98\%\notag
	\end{align}
\item Legal reasoning
	\par\smallskip
	Suppose a crime has been committed. Blood is found at the scene for which there is no innocent explanation.
	It is of a type which is present in \(1\%\) of the population.
	\begin{exercise}
	\item
		The prosecutor claims: "There is a \(1\%\) chance that the defendant would have the crime blood type if he
		were innocent. Thus there is a \(99\%\) chance he is guilty." This is known as the prosecutor's fallacy.
		What is wrong with this argument?
		\par\smallskip
		We are given that the probability of prevalence is \(P(H = 1) = 0.01\) so \(P(H = 0) = 0.99\). Moreover, the
		the number of people in the community with this rare blood type is \(8,000\) and there is no innocent
		explanation for it. That is, the true negative rate, or the probability that person doesn't have this blood
		and is not guilty is \(1\).
		\begin{table}
		\centering
			\begin{tabular}{l | c c}
				& Y = 0 & Y = 1\\ \hline
				H = 0 & 1 & 0\\
				H = 1 & 7999 / 8000 & 1 / 8000
			\end{tabular}
			\caption{The probability matrix for guilty and rare blood and vice versa.}
			\label{ch2_ex21}
		\end{table}
		From the probability matrix in \cref{ch2_ex21}, we have that true negative rate is \(100\%\) and the true 
		positive rate is \(1 / 8000\) for the given blood type; therefore, the probability that we are guilty with the
		rate blood type is
		\begin{align}
			p(H = 1\mid Y = 1) 
			&= \frac{p(Y = 1\mid H = 1) p(H = 1)}{p(Y = 1\mid H = 1) p(H = 1) + p(Y = 1\mid H = 0) p(H = 0)}
			\notag\\
			&= \frac{1 / 8000\cdot 1 / 100}{1 / 8000\cdot 1 / 100 + 7999 / 8000 \cdot 99 / 100}\notag\\
			&\approx 0.000126\%\notag
		\end{align}
	\item
		The defender claims: "The crime occurred in a city of \(800,000\) people. The blood type would found in
		approximately \(8,000\) people. The evidence has provided a probability of just \(1\) in \(8,000\) that the
		defendant is guilty, and thus has no relevance." This is known as the defender's fallacy. 
		What is wrong with this argument?
		\par\smallskip
		As for the defenders argument, the probability the defendant has the rare blood and is not guilty is
		\begin{align}
			p(H = 1\mid Y = 0) 
			&= \frac{p(Y = 0\mid H = 1) p(H = 1)}{p(Y = 0\mid H = 1) p(H = 1) + p(Y = 0\mid H = 0) p(H = 0)}
			\notag\\
			&= \frac{7999 / 8000\cdot 1 / 100}{7999 / 8000\cdot 1 / 100 + 1 \cdot 99 / 100}\notag\\
			&\approx 1\%\notag
		\end{align}
	That is, \(p(H = 1\mid Y = 0) \approx 1\% > 1 / 8000 = 0.0125\%\).
	\end{exercise}
\item Probabilities are sensitive to the form of the question that was used to generate the answer
	\par\smallskip
	My neighbor has two children. Assuming that the gender of a child is like a coin flip, it is most likely, 
	\textit{a priori}, that my neighbor has one boy and one girl, with probability \(1 / 2\). The other possibilities--two
	boys or two girls--have probabilities \(1 / 4\), respectively.
	\begin{exercise}
	\item
		Suppose I ask him whether he has any boys, and he says yes. What is the probability that one child is a 
		girl?
	\item
		Suppose instead that I happen to see one of his children run by, and it is a boy. What is the probability that
		the other child is a girl?
	\end{exercise}
\item Deriving the posterior predictive density for the healthy levels game
	\par\smallskip
	W will first consider one-dimensional "rectangles" (i.e., lines); since the dimensions are independent, we can
	easily generalize to 2d.
	\par\smallskip
	For convenience, we will follow the notation of Josh Tenenbaum's Ph.D. thesis. In particular, let \(h = \theta\) be
	the unknown hypothesis or parameter vector, \(\mathcal{H}\) be the set of possible hypotheses (rectangles), 
	\(\mathcal{H}_y\) be the set of hypotheses consistent with observation \(y\) (so the rectangles have to be big 
	enough to capture \(y\)), and \(\mathcal{H}_{\mathcal{D}, y}\) be the set of hypotheses consistent with all the
	examples in \(\mathcal{D}\) as well as with \(y\).
	\par\smallskip
	The posterior predictive is given by \(p(y\mid\mathcal{D}) = p(y\mid\mathcal{D}) \ p(\mathcal{D})\), where
	\begin{align}
		p(\mathcal{D}) &= \int_{h\in\mathcal{H}}p(h)p(\mathcal{D}\mid h)dh\\
		&= \int_{h\in\mathcal{H_{\mathcal{D}}}}\frac{p(h)}{\lvert h\rvert^N}dh\label{ch2_eq1}
	\end{align}
	where we used the fact that \(p(\mathcal{D}\mid h) = 1 / \lvert h\rvert^N\) if \(h\in\mathcal{H_{\mathcal{D}}}\)
	and is \(0\) otherwise. Similarly, 
	\(p(y, \mathcal{D}) = \int_{h\in\mathcal{H}_{\mathcal{D}, y}}p(h) / \lvert h\rvert^Ndh\).
	\par\smallskip
	To derive the integral in \cref{ch2_eq1}, let us assume the maximum observed value is \(0\) (we can pick any
	maximum and recenter the data, since we assume a translation invariant prior). Then the right edge of the
	rectangle must lie past the data, so \(\ell\geq 0\). Also, if \(r\) is the range spanned by the examples, then the
	left most data point is \(-r\), so the left side of the rectangle must satisfy \(\ell - s\leq -r\), where \(s\) is the size
	of the rectangle.
	\begin{exercise}
	\item
		Using these assumptions, show that
		\[
			p(\mathcal{D}) = \frac{1}{N(N - 1)r^{N - 1}}
		\]
		Hint: use integration by parts
		\[
			I = \int_a^bf(x)g'(x)dx = f(x)g(x)\rvert_a^b - \int_a^bf'(x)g(x)dx
		\]
	\item
		To compute \(p(y, \mathcal{D})\), we just need to extend the range from \(r\) to \(r + d\), where \(d\) is the
		distance from \(y\) to the closest observed example. Hence show that
		\[
			p(y\mid\mathcal{D}) = \frac{1}{(1 + d/r)^{N - 1}}
		\]
	\end{exercise}
\end{exercise}

